---
title: "TAREA 3"
author: " Jovanna Ramírez Guerrero"
date: "2025-04-13"
output: 
 html_document:
    toc: true
    toc_depth: 1
    theme: cerulean
---

<style>
body {
text-align: justify}
</style>

<style type="text/css">
body{/*Normal*/
     font-size: 20px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Desarrolla, a partir de los visto en clase, el ejercicio 5 del Capítulo 3. Toma como referencia el archivo <b>Ejemplos/Ejemplo_Capítulo_3.Rmd</b>.

Consider the following utility functions for two alternatives, namely i and j:

<b>$U_i = V_i + ε_i$</b>


<b>$U_j = V_j + ε_j$</b>

Assume that the difference between the error terms below follows the triangle distribution:

<b>$ε_q=ε_i-ε_j$</b>

Parting from the assumption above, derive a binary choice model for the probability of selecting alternative j.

```{r}
f_den <- function(x, B=1){
  if(x>=-B & x<=0){
    y <- 1/B^2 *x +1/B
  }else if(x>0 & x<=B){
    y <- -1/B^2 *x +1/B
}else{
    y <- 0
}
  return(y)
}

fv_den <- Vectorize(f_den, "x")

f_dis <- function(x, B=1){
  if(x<(-1)*B){
    y <- 0}else if(x>=-B & x<=0){
    y <- (x+B)*(fv_den(x)/2)
  }else if(x>0 & x<=B){
    y <- 1- (B-x)*fv_den(x)/2
} else{
    y <- 0
}
  return(y)
}

x <- seq(-1.5, 1.5, length.out=500)
y <- sapply(x, f_den)

ggplot()+
  geom_line(aes(x,y), color="black", , linewidth=1.5)+
  geom_function(fun = fv_den, color="deeppink4", linewidth=0.5)+
  xlim(-1.5, 1.5)+
  theme_bw()

decision <- function(Vj, Vi, B=1){
  Vji <- Vj-Vi
  PVj <- f_dis(Vji)
  
  ggplot()+
  geom_function(fun = fv_den, color="black", linewidth=0.5, xlim=c(-B-0.1, B+0.1))+
  stat_function(fun = fv_den, geom= "area", fill="deeppink4", alpha=0.5, xlim=c(-B, Vji))+ 
  geom_text(aes(B/2, 1/(2*B), label=paste0("Prob(Vj)=", round(PVj,2), sep="")))+
    labs(title="Probabilidad de elección")+
  theme_bw()
  
}

decision(Vj=6, Vi=5.5)
```

La probabilidad de seleccionar la alternativa j es de $P(V_j)=0.88$.
