---
title: "TAREA 2"
author: " Jovanna Ramírez Guerrero"
date: "2025-04-13"
output: 
  html_document:
    toc: true
    toc_depth: 1
    theme: cerulean
---

<style>
body {
text-align: justify}
</style>

<style type="text/css">
body{/*Normal*/
     font-size: 20px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(discrtr) # A companion package for the book Introduction to Discrete Choice Analysis with `R`
library(mlogit) # Multinomial Logit Models
library(readr) # Read Rectangular Text Data

library(gplots) # Various R Programming Tools for Plotting Data
library(ggmosaic) # Mosaic Plots in the 'ggplot2' Framework
library(treemapify)
library(ggridges)
library(ggalluvial)
library(plotly)
```

A partir del archivo <b>Ejercicios/Capítulo_2.Rmd</b> explora el conjunto de datos <b>Car</b> y realiza los ejercicios del capítulo 2 <b>Exploratory Data Analysis</b>.

Call data set Car from package mlogit. Once you have loaded the data set, do the following:

<ul>
 <li>How many variables are there in this data set and of which type (i.e., categorical/quantitative)?</li>

 <li>Choose four relevant categorical variables from this data set and describe them using univariate, bivariate, and multivariate techniques. Discuss your results.</li>

 <li>What ideas about individuals’ choices regarding car do you develop from this EDA exercise?</li>
</ul>

Se carga el conjunto de datos, se muestran las variables y su tipo:

```{r}
data(Car)
glimpse(Car)
```

En este caso es necesario añadir columnas acerca de la elección, las gráficas se realizarán con la data frame `new_car`:

```{r}
new_car <- Car
choice_char <- gsub("choice","", new_car$choice)
new_car <- new_car |> mutate(choice_char = choice_char)

type <- vector("character", nrow(new_car))
fuel <- vector("character", nrow(new_car))
price <- vector("numeric", nrow(new_car))
range <- vector("numeric", nrow(new_car))
acc <- vector("numeric", nrow(new_car))
speed <- vector("numeric", nrow(new_car))
pollution <- vector("numeric", nrow(new_car))
size <- vector("numeric", nrow(new_car))
space <- vector("numeric", nrow(new_car))
cost <- vector("numeric", nrow(new_car))
station <- vector("numeric", nrow(new_car))
for (i in 1:nrow(new_car)) {
  type[i] <- as.character(new_car[i, paste0("type", new_car$choice_char[i], collapse = "")])
  fuel[i] <- as.character(new_car[i, paste0("fuel", new_car$choice_char[i], collapse = "")])
  price[i] <- new_car[i, paste0("price", new_car$choice_char[i], collapse = "")]
  range[i] <- new_car[i, paste0("range", new_car$choice_char[i], collapse = "")]
  acc[i] <- new_car[i, paste0("acc", new_car$choice_char[i], collapse = "")]
  speed[i] <- new_car[i, paste0("speed", new_car$choice_char[i], collapse = "")]
  pollution[i] <- new_car[i, paste0("pollution", new_car$choice_char[i], collapse = "")]
  size[i] <- new_car[i, paste0("size", new_car$choice_char[i], collapse = "")]
  space[i] <- new_car[i, paste0("space", new_car$choice_char[i], collapse = "")]
  cost[i] <- new_car[i, paste0("cost", new_car$choice_char[i], collapse = "")]
  station[i] <- new_car[i, paste0("station", new_car$choice_char[i], collapse = "")]
}

new_car <- new_car |> mutate(choice_num= factor(choice_char), type=type, fuel=fuel, price=price, range=range,
                             acc=acc, speed=speed, pollution=pollution, size=size, space=space, cost=cost,
                             station=station)

new_car$type <- factor(new_car$type, levels = levels(new_car$type1))
new_car$fuel <- factor(new_car$fuel, levels = levels(new_car$fuel1))
```

Análisis Univariado

```{r}
new_car <- new_car %>%
  mutate(price_cat = cut(price,
                         breaks = quantile(price, probs = c(0, 0.33, 0.66, 1), na.rm = TRUE),
                         labels = c("Bajo", "Medio", "Alto"),
                         include.lowest = TRUE))

ggplot(new_car, aes(x = price_cat)) +
  geom_bar(fill = "cyan") +
  labs(title = "Distribución por nivel de precio", x = "Precio", y = "Frecuencia")

ggplot(new_car, aes(x = type)) +
  geom_bar(fill = "pink") +
  labs(title = "Distribución por tipo de auto", x = "Tipo", y = "Frecuencia")

ggplot(new_car, aes(x = fuel)) +
  geom_bar(fill = "darkseagreen") +
  labs(title = "Distribución por tipo de combustible", x = "Combustible", y = "Frecuencia")

new_car <- new_car %>%
  mutate(cost_cat = cut(cost,
                        breaks = quantile(cost, probs = c(0, 0.33, 0.66, 1), na.rm = TRUE),
                        labels = c("Bajo", "Medio", "Alto"),
                        include.lowest = TRUE))

ggplot(new_car, aes(x = cost_cat)) +
  geom_bar(fill = "orange") +
  labs(title = "Distribución por nivel de costo de uso", x = "Costo", y = "Frecuencia")

```

Análisis bivariado

```{r}
# price_cat vs fuel

ggplot(new_car, aes(x = fuel, fill = price_cat)) +
  geom_bar(position = "fill") +
  labs(title = "Nivel de precio según tipo de combustible", y = "Proporción")

# price_cat vs type

ggplot(new_car, aes(x = type, fill = price_cat)) +
  geom_bar(position = "fill") +
  labs(title = "Nivel de precio según tipo de auto", y = "Proporción")

# price_cat vs cost_cat

ggplot(new_car, aes(x = cost_cat, fill = price_cat)) +
  geom_bar(position = "fill") +
  labs(title = "Relación entre costo de uso y precio", x = "Costo", y = "Proporción del precio")

```

Análisis Multivariado

```{r}
# type vs fuel vs price_cat

ggplot(new_car, aes(x = type, fill = price_cat)) +
  geom_bar(position = "fill") +
  facet_wrap(~fuel) +
  labs(title = "Tipo de auto y precio según combustible")

# fuel vs cost_cat vs price_cat

ggplot(new_car, aes(x = cost_cat, fill = price_cat)) +
  geom_bar(position = "fill") +
  facet_wrap(~fuel) +
  labs(title = "Precio según tipo de combustible y costo de uso")

```

